---
title: "Warfarin Dosing Service Analysis"
subtitle: "Exploratory Analysis"
author: "Brian Gulbis"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
    html_notebook:
      code_folding: hide
      fig_caption: yes
      toc: yes
      toc_float: yes
---

```{r library, message=FALSE}
library(BGTools)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(scales)
library(tableone)
```

```{r data}
tmp <- get_rds("tidy")
tmp <- get_rds("analysis")
pts.include <- inner_join(pts.include, data.new, by = "pie.id")
```

## Warfarin Utilization

```{r warfarin_use, fig.cap="Number of Patients Receiving Warfarin", fig.height=6, fig.width=6}
d <- analyze.patients.all %>%
    filter(year.cal >= 2013) 

g <- ggplot(d, aes(x = year.cal)) +
    geom_bar() +
    ggtitle("Number of Patients Receiving Warfarin") +
    xlab("Year") +
    ylab("Number of Patients") +
    scale_y_continuous(limits = c(0, 3000))

print(g)
```

```{r services, fig.cap="Top 10 Medical Services Utilizing Warfarin", fig.height=6, fig.width=6}
d <- analyze.services.all %>%
    mutate(service = ordered(service, levels = names(sort(table(service), decreasing = TRUE)))) %>%
    group_by(service) %>%
    summarize(count = n()) %>%
    arrange(desc(count)) %>%
    top_n(10, count)

g <- ggplot(d, aes(x = service, y = count)) +
    geom_bar(stat = "identity") +
    ggtitle("Top 10 Medical Services Utilizing Warfarin") +
    xlab("Medical Service") +
    ylab("Number of Patients") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(g)
```

#### All Time

```{r ds_by_service, fig.cap="Warfarin Dosing Service Utilization Among Top 10 Medical Services Ordering Warfarin"}
d <- tmp <- analyze.services.all %>%
    mutate(service = ordered(service, levels = names(sort(table(service), decreasing = TRUE)))) %>%
    group_by(service, consult, historical) %>%
    summarize(count = n()) %>%
    arrange(desc(count)) %>%
    filter(service <="Orthopedic Surgery Service") %>%
    group_by(service, historical) %>%
    spread(consult, count) %>%
    mutate(`FALSE` = `FALSE` - `TRUE`)

d$`FALSE`[d$`FALSE` < 0] <- 0

d <- gather(d, consult, count, -service, -historical)

g <- ggplot(d, aes(x = service, y = count, fill = consult)) +
    geom_bar(stat = "identity") +
    ggtitle("Warfarin Dosing Service Utilization Among\nTop 10 Medical Services Ordering Warfarin") +
    xlab("Medical Service") +
    ylab("Number of Patients") +
    scale_fill_brewer(palette = "Set1") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(g)
```

#### Current vs. Historical

```{r}
print(g + facet_grid(historical ~ .))
```

```{r hospital_units, fig.cap="Top 10 Hospital Units for Warfarin Initiation", fig.height=6, fig.width=6}
d <- analyze.locations.all %>%
    mutate(location = ordered(location, levels = names(sort(table(location), decreasing = TRUE)))) %>%
    group_by(location) %>%
    summarize(count = n()) %>%
    arrange(desc(count)) %>%
    top_n(10, count)

g <- ggplot(d, aes(x = location, y = count)) +
    geom_bar(stat = "identity") +
    ggtitle("Top 10 Hospital Units for Warfarin Initiation") +
    xlab("Hospital Unit") +
    ylab("Number of Patients") +
    scale_fill_brewer(palette = "Set1") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(g)
```


## Warfarin Dosing Service Utilization

```{r service_utilization, fig.cap="Daily Warfarin Orders and Warfarin Consults", fig.height=6, fig.width=6}
g <- ggplot(data = analyze.utilization, aes(x = action.date, y = n, color = order)) +
    geom_line(size = 0.2, alpha = 0.7) +
    geom_smooth() + 
    # geom_smooth(method = "lm", linetype = "dashed") +
    ggtitle("Daily Warfarin Orders and Warfarin Consults") +
    xlab("Date") + 
    ylab("Number of Daily Orders") +
    scale_color_brewer(palette = "Set1") +
    xlim(c(mdy_hm("1/1/2013 00:00"), mdy_hm("12/31/2015 23:59"))) 
    # coord_cartesian(xlim = c(mdy_hm("1/1/2013 00:00"), mdy_hm("12/31/2015 23:59"))) 

print(g)
```

```{r percent_orders, fig.cap="Percent of Daily Warfarin Orders Managed by Pharmacy Dosing Service", fig.height=6, fig.width=6}
d <- group_by(analyze.utilization, action.date) %>%
    spread(order, n) %>%
    # rename(warfarin = `FALSE`, consults = `TRUE`) %>%
    mutate(perc.consults = consult / warfarin) 

g <- ggplot(d, aes(x = action.date, y = perc.consults)) +
    geom_line(size = 0.2, alpha = 0.7) +
    geom_smooth() +
    # geom_smooth(method = "lm", color = "red") +
    ggtitle("Percent of Daily Warfarin Orders Managed by Pharmacy Dosing Service") +
    xlab("Date") +
    ylab("Orders (%)") +
    xlim(c(mdy_hm("1/1/2013 00:00"), mdy_hm("12/31/2015 23:59"))) 

print(g)
```

## Comparison

### Groups

* Current: warfarin started between January 1, 2015 and December 31, 2015
* Historical: warfarin started between July 1, 2012 and December 31, 2014

```{r groups, results='asis'}
total <- pts.include %>%
    group_by(group, year) %>%
    summarize(total = n())

groups <- total %>%
    spread(year, total)

knitr::kable(groups, caption = "Number of patients per group")
```

### Demographics

#### All Patients
```{r demographics, results='asis'}
demograph <- pts.include %>% 
    inner_join(data.demographics, by = "pie.id") %>%
    inner_join(data.measures, by = "pie.id") %>%
    mutate(sex = factor(sex, levels = c("Female", "Male")),
           race = str_replace_all(race, "Oriental", "Asian"),
           race = factor(race, exclude = c("", "Unknown")))

vars <- c("age", "sex", "bmi", "race", "length.stay", "therapy")
fvars <- c("sex", "race", "therapy")
tbl <- CreateTableOne(vars, "group", demograph, fvars)
ptbl <- print(tbl, nonnormal = c("age", "bmi", "length.stay"), printToggle = FALSE)
knitr::kable(ptbl, caption = "Demographics for Current Year")
```

#### Current
```{r demograph_curr, results='asis'}
d <- filter(demograph, year == "current")
tbl <- CreateTableOne(vars, "group", d, fvars)
ptbl <- print(tbl, nonnormal = c("age", "bmi", "length.stay"), printToggle = FALSE)
knitr::kable(ptbl, caption = "Demographics for Historical Years")
```

#### Historical
```{r demograph_hist, results='asis'}
d <- filter(demograph, year == "historical")
tbl <- CreateTableOne(vars, "group", d, fvars)
ptbl <- print(tbl, nonnormal = c("age", "bmi", "length.stay"), printToggle = FALSE)
knitr::kable(ptbl, caption = "Demographics for Historical Years")
```

### Indications

#### All Patients
```{r indications, fig.cap="Indications for warfarin use", fig.height=6, fig.width=6}
d <- pts.include %>%
    inner_join(data.warfarin.indications, by = "pie.id") %>%
    gather(indication, value, -pie.id, -group, -year) %>%
    filter(value == TRUE) %>%
    group_by(group, year, indication) %>%
    summarize(n = n()) %>%
    left_join(total, by = c("group", "year")) %>%
    mutate(perc = n / total) 

g <- ggplot(d, aes(x = indication, fill = group)) +
    geom_bar(aes(y = perc), stat = "identity", position = "dodge") + 
    facet_grid(year ~ .) +
    ggtitle("Indications for warfarin use") +
    ylab("Patients (%)") +
    scale_y_continuous(labels = percent) +
    scale_fill_brewer(palette = "Set1") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(g)
```


### Dispositions

```{r disposition, fig.cap="Disposition on Discharge", fig.height=6, fig.width=6}
d <- pts.include %>% 
    inner_join(data.demographics, by = "pie.id") %>%
    mutate(disposition = str_replace_all(disposition, regex(".*expired.*|.*hospice.*|.*deceased.*", ignore_case = TRUE), "Deceased/Hospice"),
           disposition = str_replace_all(disposition, regex(".*home.*|.*against.*", ignore_case = TRUE), "Home"),
           disposition = str_replace_all(disposition, regex(".*dc.*|.*transfer.*|.*care.*|.*skill.*", ignore_case = TRUE), "Transferred"),
           disposition = factor(disposition)) %>%
    group_by(group, year, disposition) %>%
    summarize(n = n()) %>%
    left_join(total, by = c("group", "year")) %>%
    mutate(perc = n / total) 
    
g <- ggplot(d, aes(x = disposition, fill = group)) +
    geom_bar(aes(y = perc), stat = "identity", position = "dodge") + 
    facet_grid(year ~ .) +
    ggtitle("Disposition on Discharge") +
    ylab("Patients (%)") +
    scale_y_continuous(labels = percent) +
    scale_fill_brewer(palette = "Set1") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(g)
```


## Exploratory Plots

### Change in INR

#### All Patients

```{r inr, fig.cap="Change in INR after starting warfarin"}
d <- pts.include %>%
    inner_join(data.labs.inrs, by = "pie.id") %>%
    calc_lab_runtime(units = "days")

g <- ggplot(d, aes(x = run.time, y = lab.result, group = group)) +
    geom_point(alpha = 0.3, size = 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "green") +
    geom_smooth(aes(color = group)) + 
    ggtitle("Change in INR after starting warfarin") +
    xlab("Day") +
    ylab("INR") +
    scale_x_continuous(breaks = seq(-2, 10, by = 2)) +
    scale_color_brewer(palette = "Set1") +
    coord_cartesian(xlim = c(-2, 10), ylim = c(1, 4)) 

print(g)
```

#### Groups

```{r}
print(g + facet_grid(year ~ therapy))
```

### Time in Therapuetic Range

#### All Patients
```{r inrange, fig.cap="Percent of time INR is within therapeutic range"}
d <- inner_join(pts.include, data.inr.inrange, by = "pie.id") 

g <- ggplot(d, aes(x = group, y = perc.time)) +
    geom_boxplot() +
    ggtitle("Percent of time INR is within therapeutic range") +
    ylab("Time in Therapeutic Range (%)")

print(g)
```

#### Groups

```{r}
print(g + facet_grid(year ~ therapy))
```


### Critical INR Values

#### All Patients
```{r critical_inr, fig.cap="Percent of time INR is critical (above 4)"}
d <- inner_join(pts.include, data.inr.supratx, by = "pie.id") 

g <- ggplot(d, aes(x = group, y = perc.time)) +
    geom_boxplot() +
    ggtitle("Percent of time INR is critical (above 4)") +
    ylab("Time Supratherapeutic (%)")

print(g)
```

#### Groups

```{r}
print(g + facet_grid(year ~ therapy))
```

### Change in Hgb

#### All Patients
```{r hgb, fig.cap="Change in hemoglobin after starting warfarin"}
d <- pts.include %>%
    inner_join(data.labs.hgb, by = "pie.id") %>%
    calc_lab_runtime(units = "days") 

g <- ggplot(d, aes(x = run.time, y = lab.result, group = group)) +
    geom_point(alpha = 0.3, size = 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "dark blue") +
    geom_smooth(aes(color = group)) + 
    ggtitle("Change in hemoglobin after starting warfarin") +
    xlab("Day") +
    ylab("Hgb (g/dL)") +
    scale_x_continuous(breaks = seq(-2, 10, by = 2)) +
    scale_color_brewer(palette = "Set1") +
    coord_cartesian(xlim = c(-2, 10), ylim = c(5, 15))

print(g)
```

#### Groups
```{r hgb2, fig.cap="Change in hemoglobin after starting warfarin"}
print(g + facet_grid(year ~ therapy))
```
